---
title: "Files preparation and importing to `phyloseq`"
author: "Luis-Vargas, Maira N."
output: github_document
always_allow_html: true
---

email: nayeli.luis@ciencias.unam.mx

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is `phyloseq`?

`phyloseq` is a R package which brings tools in order to analyse microbial communities from many types of data. The Github  of `phyloseq` ([click here](https://joey711.github.io/phyloseq/), has many tutorials for make several analysis and data visualization. But, in my opinion, is unclear how the data must be imported, in this tutorial I hope to show an easy way to import data for using `phyloseq`. 

## First of all

We need four files, three of these files are insde `.qza` directories from QIIME2, the last one is your `sample-metadata.csv' file. The next table shows the directories where the files are, you just have to unzip them.

```{r, include=TRUE, echo=FALSE, message=FALSE}

library(kableExtra)

files <- data.frame(
  Directory = c("table.qza", "unrooted-tree.qza", "taxonomy.qza", ""),
  Files = c("feature-table.biom", "tree.nwk", "taxonomy.tsv", "sample-metadata. csv"))

kbl(files, "latex" ,caption = "Necessary files",
    align = c("c", "c"),
    escape = FALSE, 
    booktabs = TRUE,
    linesep = "") %>%
  kable_paper("hover", full_width = T)
```

From these file, we will create others, which are the ones that will be import into `phyloseq`. These are: `feature-table.tsv` and `otu_tax_matrix.csv`. 

## The preparation

### `feature-table.tsv`

To get the `feature-table.tsv` file we need to transform the `biom` file. If you want to know what a biom file is, yo can go to the [Biom page](https://biom-format.org). In this case we have to transform the `biom` file in a `.tsv` file. For that you'll active your QIIME2 and copy-paste the next comand. 

`biom convert -i feature-table.biom -o feature-table.tsv --to-tsv`


The `.tsv` file looks like this:

```{r, message=FALSE}
library(tidyverse)
library(magrittr)

asv_table <- read.table("../datasets/feature-table.tsv", sep = "\t")
head(asv_table)
```

The first column is the FeatureID (ASV) and the next ones are the samples. So, the next step is put column names and we can do this importing the `sample-metadata.csv` file, passing the names of the samples as columnames. 

```{r}

metadata <- read.csv("../datasets/sample-metadata.csv", row.names = 1)
sample.names <- rownames(metadata)

colnames(asv_table) <- c("Feature.ID", sample.names)

head(asv_table)

```

Our first file (`feature-table.tsv`) is ready!


### `otu_tax_matrix.csv`. 

This file has the OTUIDs, the taxonomy and the samples, and the taxonomy should be separated. I mean, each taxonomic hierarchy must be in a column. To get this file we need the `taxonomy.tsv` file and the `feature-table.tsv`.  

First, the `taxonomy.tsv` has three columns: `Feature.ID`, `Taxon` and `Confidence`. 

```{r}

tax <- read.table("../datasets/taxonomy.tsv", sep = '\t', header = T)
colnames(tax)

```

The first step is delete the column `Confidence`. Next, we'll rename the column  `Feature.ID` in `OTUID`, because we will merge the datasets by this column. And finally, we'll separate the column `Taxon` in different columns. 

```{r, message=FALSE, warning=FALSE}
tax %<>% 
  select(-Confidence) %>%
  separate(Taxon, c("Kingdom", "Phylum"), sep = "; p__") %>%
  separate(Phylum, c("Phylum", "Class"), sep = "; c__") %>%
  separate(Class, c("Class", "Order"), sep = "; o__") %>%
  separate(Order, c("Order", "Family"), sep = "; f__") %>%
  separate(Family, c("Family", "Genus"), sep = "; g__") %>%
  separate(Genus, c("Genus", "Specie"), sep = "; s__") %>%
  mutate_all(na_if,"")

tax$Kingdom <- sub (pattern = "d__", replacement = "", 
                       tax$Kingdom)

head(tax)
```

And, we'll merge the datasets. In order to have all de taxonomy with the number of seqs by ASV and by sample.

```{r}
asv_matrix <- merge(x = asv_table, y = tax, by = "Feature.ID")
head(asv_matrix)
```

## Importing datasets to `phyloseq`

Finally, we can import our data to `phyloseq`:

```{r}
library(phyloseq)

# Column Feature.ID as rownames in the asv_table
rownames(asv_table) <- asv_table$Feature.ID
asv_table %<>% select(-Feature.ID)

# Column Feature.ID as rownames in the df taxonomy
rownames(tax) <- tax$Feature.ID
tax %<>%
  select(-Feature.ID) %>%
   as.matrix(.) # Convert to a matrix


ASV <- otu_table(asv_table, taxa_are_rows = TRUE)
TAX <- tax_table(tax)
metadata <- sample_data(metadata)
phy_tree <- read_tree("../datasets/tree.nwk")

phyloseq_object <- phyloseq(ASV, TAX, metadata, phy_tree)
phyloseq_object
```

I recommend you to save this physeq object as a RDS

```{r}
saveRDS(phyloseq_object, "../datasets/phyloseq_object.RDS")
```

