---
title: "Rarefy clean data and plot"
author: "Luis-Vargas, Maira N."
output: github_document
---

email: nayeli.luis@ciencias.unam.mx

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Resource
All the functions are contained in `funs_rarefy.R` in `scripts` directory. 

```{r}
source("../scripts/funs_rarefy2.R")
```


## Dataset preparation

The dataset that QIIME2 returns is called `observed_features.csv` and it was obtained with the `qiime diversity alpha-rarefaction plugin`. 
Usually, the dataset has a column for samples names (`sample_id`) at the begining and other categorical columns like `description`, `site`, etc. at the end it depends on your project.  

The dataset looks like this:

```{r, include=FALSE}
obs <- read.csv("../datasets/observed_features.csv")
head(obs)
```


## First step: Data cleaning

In order to do this we will use the function `clean_data()`. this function reformat the dataset and separate the columns by `depth_seqs()`, `iter` and `asvs`.



```{r, echo=TRUE, include = TRUE, warning=FALSE}
obs_long <- clean_data(obs)
head(obs_long)
```


## Second step: Get the mean of observed ASVs

The second step is to obtain the mean aof the number of ASVs observerd in each iteration at each depth. first, we have to nest our dataset in function of a variable, could be a categorical variable or the name of our samples.

```{r, echo=TRUE, include = TRUE, warning=FALSE}
obs_nest <- obs_long %>% 
  group_by(depth_seqs, site) %>% 
  nest()
head(obs_nest)
```

To check the kind of object `obs_nest`:

```{r}
class(obs_nest)
```

To acces data:

```{r}
obs_nest$data[[1]]
```

Finally, we can get the mean of observed ASVs with `asvs_mean()` fuction.

```{r}
obs_mean <- mutate(
  obs_nest, mean_asv = map_dbl(.x = data, .f = asvs_mean)
) %>% 
  select(-data)

# Convert depth_seqs in a numeric variable
obs_mean$depth_seqs <- sapply(obs_mean$depth_seqs, as.numeric)

head(obs_mean)
```


## Third step: Plot

Finally just plot the last object. 

```{r}
obs_mean %>% 
  ggplot(aes(x = depth_seqs,
             y = mean_asv, 
             color = site)) +
  geom_line() +
  geom_point(alpha = 0.5) +
  theme_bw() 
```

## To plot each sample

```{r}
# Group by sample_id
obs_sample <- obs_long %>% 
  group_by(depth_seqs, sample_id) %>% 
  nest()

# Get mean of asvs by sample
obs_mean_sample <- mutate(
  obs_sample, mean_asv = map_dbl(.x = data, .f = asvs_mean)
) %>% 
  select(-data)

# Convert depth_seqs in a numeric variable
obs_mean_sample$depth_seqs <- sapply(obs_mean_sample$depth_seqs, as.numeric)

# Check dataset
head(obs_mean_sample)
```
```{r}
# plot
obs_mean_sample %>% 
  ggplot(aes(x = depth_seqs,
             y = mean_asv, 
             color = sample_id)) +
  geom_line() +
  geom_point(alpha = 0.5) +
  theme_bw()
```

