---
title: "Venn Diagram: Shared ASVs"
author: "Luis-Vargas Maira N"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
The special library you need is `limma`. 

```{r, message=FALSE}
library(tidyverse)
library(magrittr)
library(reshape2) #to melt dataset
library(limma)
```


# Data preparation

We'll use the `otu_tax_matrix.csv` file, it was generated with `taxonomy.tsv` file  and `feature-table.tsv` file (see the tuto '[Files preparation and importing to `phyloseq`](https://github.com/Nayeli-Luis/16SrRNA_FromQIIME2_ToR/blob/main/tutorials/phyloseq-files-preparation-importing.md)).

The colnames of the `otu_tax_matrix.csv` looks like this: 

```{r, message=FALSE, warning=FALSE}
otu.tax <- read_csv('../datasets/otu_tax_matrix.csv')

colnames(otu.tax)

```

It contains the `OTUID` variable with the ASV identifier, your samples (in this case from `A1` to `D10`) with the counts of each ASV and a column by each taxonomic hierarchy (from `Kingdom` to `Specie`). 

In this `otu_tax_matrix.csv` we have four sites  (A, B, C and D) with 10 repetitions. We want to **know how many ASVs shared between them**. 

Let's check dimension. 

```{r}
dim(otu.tax)
```


## Step 1. Copy-paste
Save a copy of `otu.tax`, because we will use it to search the taxonomic classification of the ASVs shared.

```{r}
otu.tax.copy <- otu.tax
```

## Step 2. Transform to long format
Transform the dataset to  long format with `melt()`. 

```{r}
otu.tax %<>% melt() 
colnames(otu.tax)
```

You will have two new columns `variable` and `value`, `variable` is the name of the samples and `value` the count of your ASVs. 

Let's check dimension, as you can see you have more observations. 
```{r}
dim(otu.tax)
```

## Stept 3. Obtain sum of ASVs by site. 

At this moment, we don't need each repetition of site, because we want to compare between our four sites (not betwenn our 40 samples). So, we'll add a new variable called `site`, obviously it will be a category. It is easy with the function `case_when()` and `grepl()`

```{r}
# Create new variable

otu.tax %<>% 
  mutate(site = case_when(
    grepl("A", variable) ~ "A",
    grepl("B", variable) ~ "B",
    grepl("C", variable) ~ "C",
    grepl("D", variable) ~ "D"
  ))
```

Now we'll group by site and get the mean of ASVs of each site. 

```{r, message=FALSE, warning=FALSE}
otu.tax %<>%
  group_by(OTUID, site) %>%
  summarise(across(value, .fns = list(sum = sum)))

head(otu.tax)
```

## Step 4. Tranform to wide format

```{r}
otu.tax %<>% 
  pivot_wider(names_from = "site", values_from = "value_sum")
head(otu.tax)
```

## Step 5. Generate presence/absence matrix

In this step we mark the absence of presence of an ASV with 1 and 0  by site. If the sum of ASVs is greater than or equal to 1, we'll mark 1, if not, we'll mark 0. 

```{r}
count_matrix <- otu.tax %>%
  transmute(
    OTUID,
    A = if_else(A >= 1, 1, 0), 
    B = if_else(B >= 1, 1, 0),
    C = if_else(C >= 1, 1, 0),
    D = if_else(D >= 1, 1, 0)
  )

head(count_matrix)
```


Transform the `count_matrix` dataset to a matrix. 

```{r}
count_matrix1 <- count_matrix %>%
  as.data.frame() %>%
  set_rownames(., .$OTUID) %>%
  select(-OTUID) %>%
  as.matrix()

head(count_matrix1)
```

# Venn Diagram: `limma` in action

Here `limma` checks the categories (in this case sites) whose shared ones, in `count_matrix`. 

```{r}
venn_counts <- vennCounts(count_matrix1)
venn_counts
```

With this data we can generate the Venn Diagram.

```{r}
my.colors<- c("#CF587E", "#E98571", "#EDB479", "#E9E19C")
vennDiagram(venn_counts, circle.col = my.colors)
```

We can see that all the sites shared 142 ASVs... Who are them?

# Taxa classification of shared ASVs

Time to use our copy of `otu.tax` and `count_matrix`. Of `otu.tax.copy` we just need the taxa.

```{r}
otu.tax.copy %<>%
  select(OTUID, Kingdom:Specie)
head(otu.tax.copy)
```

We want to know the taxonomic classification of the ASVs that are present in all sites, I mean, the ASVs with 1 in all the sites.

So, first, we'll create a vector with that ASVs. 
```{r}
asv_interest <- count_matrix %>%
  filter(A == 1 & B == 1 & C == 1 & D == 1) %>%
  .$OTUID

head(asv_interest)
```

Check if there are 142.

```{r}
length(asv_interest)
```


Then we'll search that OTUIDs in `otu.tax.copy`. 

```{r}
tax_asv <- filter(otu.tax.copy, OTUID %in% asv_interest)

head(tax_asv)
```

And that's all, continue crying with your analysis. Hope it helps you. 